### 2017-10-12 writing/paper/sections/coincidences.tex
More changes

Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
\ie, \(\av{\lambda} = \av{\gamma}\). This is equivalent to \(\lambad\) and \(\gamma\) being contained in
the same set of permutations, \ie, \(\cont{\lambda} = \cont{\gamma}\).

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a fixed
classical pattern of length 3 and \(p\) is a mesh pattern of length 2 that
varies. The classical pattern \(\pi\) will be called the \emph{dominating
pattern}. Given such a dominating pattern \(\pi\) we will write \(p_1
\coinc{\pi} p_2\) if for the mesh patterns, \(p_1\) and \(p_2\),  the sets
\(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are equal.

Our first step is to calculate whether \(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are
equal up to permutations of length 11 computationally, if they are we write
\(p_1 \coinc[comp]{\pi} p_2\). The equivalence relation \(\coinc{\pi}\) is a
refinement of the equivalence relation \(\coinc[comp]{\pi}\), and therefore the
\(\coinc{\pi}\)-equivalence classes form partitions of the
\(\coinc[comp]{\pi}\)-coincidence classes.

If \(p_1 \coinc{\pi} p_2\), it can be seen that
\(p_1 \coinc[comp]{\pi} p_2\), and therefore the \(\coinc{\pi}\)-equivalence
classes form partitions of the \(\coinc[comp]{\pi}\)-coincidence classes.

We then prove three propositions called ``Dominating Pattern Rules'', if the
First Dominating Pattern Rule shows that \(p_1 \coinc{\pi} p_2\) then we write
\(p_1 \coinc[(1)]{\pi} p_2\). If a combination of the First and Second
Dominating Pattern Rules show that \(p_1 \coinc{\pi} p_2\) then we write
\(p_1 \coinc[(2)]{\pi} p_2\); similarly we write \(p_1 \coinc[(3)]{\pi} p_2\) if
a combination of all three rules shows coincidences. These three equivalence
relations are successive refinements of \(\coinc{\pi}\). Therefore if we find
that \eg, \(\coinc[(1)]{\pi} = \coinc[comp]{\pi}\) then \(\coinc[(1)]{\pi} = \coinc{\pi}\).

@@

\begin{figure}
\begin{equation*}
p = \pattern{}{1,2}{0/1,2/2} \quad
p^\boks{2,1}= \pattern{}{1,3,2}{0/1,0/2,2/3,3/3}
\quad
p^\boksup{2,1}
= \pattern{}{1,3,2}{0/1,0/2,2/2,2/3,3/2,3/3}
\end{equation*}
\caption{The result of inserting a point into \(p=(12,\{(0,1),(2,2)\})\)}
\label{fig:addp}
\end{figure}

\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
a mesh pattern \(p^{\boks{i,j}_a} = (\tau^\prime,R^\prime)\) (\(p^{\boks{i,j}_d}\)) of length \(n+2\)
as the pattern where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\).
Formally the new underlying classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+t & \text{if } k = i+t,t\in\{1,2\}\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+2 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-2) & \text{if } \tau(k)\le j \text{ and }k> i+2\\
    \tau(k-2)+2 & \text{if } \tau(k)> j \text{ and }k> i+2\\
\end{cases}
\end{equation*}
The ordering of the top branch determines whether an ascent(or descent) is added.
The mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-2}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-2,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-2,\ell-2}\in R\}\cup\\
&\{\boks{i+1,j},\boks{i+1,j+1},\boks{i+1,j+2}\}
\end{aligned}
\end{equation*}
\end{definition}
An example of adding an ascent to a mesh pattern can be seen in \FigureRef{fig:adda}.
\begin{figure}
\begin{equation*}
p = \pattern{}{2,3,1}{0/0,1/0,1/3,
                      3/0,3/1,3/3} \quad
p^{\boks{1,1}_a}= \pattern{}{4,2,3,5,1}{0/0,
                        1/0,1/5,
                        2/0,2/1,2/2,2/3,2/5,
                        3/0,3/5,
                        5/0,5/1,5/2,5/3,5/5}
\end{equation*}
\caption{The result of inserting an ascent into \(p=(231,\{(0,0),(1,0),(1,3),(3,0),(3,1),(3,3)\})\)}
\label{fig:adda}
\end{figure}

@@

\subsection{Coincidence classes of Av(\{321, (21, \textit{R})\})}
Through experimentation, considering avoidance of permutations of up to length \(11\), we discover
that there are at least \(29\) \(\coinc[comp]{321}\)-coincidence classes where the underlying classical pattern of the mesh pattern is \(21\).

\begin{proposition}[First Dominating Pattern Rule]
    \label{prop:dom1}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 1\), then \(m_1 \coinc{\pi} m_2 \)
    as a consequence of the first dominating pattern rule if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
        \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
    \end{enumerate}
\end{proposition}

This rule can be understood in graphical form. In the pattern in
\FigureRef{fig:rule1} we can gain shading in the boxes \((0,2)\), \((2,0)\)
since if there is a point in either of these boxes there would be an occurrence of
the dominating pattern \(\perm{3,2,1}\).

@@

\begin{proof}[Proof of \PropositionRef{prop:dom1}]
    We need to prove that \(m_1 \coinc{\pi} m_2\).

    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Now we consider a permutation \(\omega \in \av{\pi}\),
    containing an occurrence of \(m_1\). If there is a point in the
    region corresponding to the box \((a,b)\), then that point, along with the
    points of the occurrence of \(m_1\), form an occurrence of \(\sigma^\boks{a,b}\).
    Then condition~\eqref{prop:dom1:cont} of the
    proposition implies an occurrence of \(\pi\). Therefore there can be no
    points in this region, which implies the occurrence of \(m_1\) is an occurrence of \(m_2\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
explained by application of \PropositionRef{prop:dom1}, thus explaining the
29 experimental coincidence classes observed.

\subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
By application of \PropositionRef{prop:dom1} we obtain \(43\) \(\coinc[(1)]{231}\)-coincidence
classes between mesh patterns with \(21\) as an underlying classical pattern.
Experimentation shows that there are \(39 \coinc[comp]{231}\)-coincidence
classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns seem to be coincident in
\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
\end{equation*}

>>>>

Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
\ie, \(\av{\lambda} = \av{\gamma}\). This is equivalent to \(\lambda\) and \(\gamma\) being contained in
the same set of permutations, \ie, \(\cont{\lambda} = \cont{\gamma}\).

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a fixed
classical pattern of length 3 and \(p\) is a mesh pattern of length 2 that
varies. The classical pattern \(\pi\) will be called the \emph{dominating
pattern}. Given such a dominating pattern \(\pi\) we will write \(p_1
\coinc{\pi} p_2\) if for the mesh patterns, \(p_1\) and \(p_2\),  the sets
\(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are equal. If this is the case we say
the two mesh patterns are \emph{coincidenc under} \(\pi\).

Our first step is to calculate whether \(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are
equal up to permutations of length 11 computationally, if they are we write
\(p_1 \coinc[comp]{\pi} p_2\). The equivalence relation \(\coinc{\pi}\) is a
refinement of the equivalence relation \(\coinc[comp]{\pi}\), and therefore the
\(\coinc{\pi}\)-equivalence classes form partitions of the
\(\coinc[comp]{\pi}\)-equivalence classes.

We then prove three propositions called ``Dominating Pattern Rules'', if the
First Dominating Pattern Rule shows that \(p_1 \coinc{\pi} p_2\) then we write
\(p_1 \coinc[(1)]{\pi} p_2\). If a combination of the First and Second
Dominating Pattern Rules show that \(p_1 \coinc{\pi} p_2\) then we write
\(p_1 \coinc[(2)]{\pi} p_2\); similarly we write \(p_1 \coinc[(3)]{\pi} p_2\) if
a combination of all three rules shows coincidence. These three equivalence
relations are successive refinements of \(\coinc{\pi}\). For future reference
we record the following:

\begin{note}
\label{note:main}
    If we find that \(\coinc[(i)]{\pi}\), for \(i = 1, 2\), or \(3\),
    equals \(\coinc[comp]{\pi}\) then the equivalence relations \(\coinc[(i)]{\pi}\), \(\coinc{\pi}\),
    and \(\coinc[comp]{\pi}\) are all equal.
\end{note}

@@

\begin{figure}
\begin{equation*}
p = \pattern{}{1,2}{0/1,2/2} \quad
p^\boks{2,1}= \pattern{}{1,3,2}{0/1,0/2,2/3,3/3}
\quad
p^\boksup{2,1}
= \pattern{}{1,3,2}{0/1,0/2,2/2,2/3,3/2,3/3}
\end{equation*}
\caption{The result of inserting a point into \(p=(12,\{\boks{0,1},\boks{2,2}\})\)}
\label{fig:addp}
\end{figure}

\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
a mesh pattern \(p^{\boks{i,j}_a} = (\tau^\prime,R^\prime)\) (\(p^{\boks{i,j}_d}\)) of length \(n+2\)
as the pattern where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\).
Formally the new underlying classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+t & \text{if } k = i+t,t\in\{1,2\}\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+2 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-2) & \text{if } \tau(k)\le j \text{ and }k> i+2\\
    \tau(k-2)+2 & \text{if } \tau(k)> j \text{ and }k> i+2\\
\end{cases}
\end{equation*}
The ordering of the top branch determines whether an ascent(or descent) is added.
The mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-2}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-2,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-2,\ell-2}\in R\}\cup\\
&\{\boks{i+1,j},\boks{i+1,j+1},\boks{i+1,j+2}\}
\end{aligned}
\end{equation*}
\end{definition}
An example of adding an ascent to a mesh pattern can be seen in \FigureRef{fig:adda}.
\begin{figure}
\begin{equation*}
p = \pattern{}{2,3,1}{0/0,1/0,1/3,
                      3/0,3/1,3/3} \quad
p^{\boks{1,1}_a}= \pattern{}{4,2,3,5,1}{0/0,
                        1/0,1/5,
                        2/0,2/1,2/2,2/3,2/5,
                        3/0,3/5,
                        5/0,5/1,5/2,5/3,5/5}
\end{equation*}
\caption{The result of inserting an ascent into \(p=(231,\{\boks{0,0},\boks{1,0},\boks{1,3},\boks{3,0},\boks{3,1},\boks{3,3}\})\)}
\label{fig:adda}
\end{figure}

@@

\subsection{Coincidence classes of Av(\{321, (21, \textit{R})\})}
Through experimentation, considering avoidance of permutations up to length \(11\), we discover
that there are \(29\) \(\coinc[comp]{321}\)-equivalence classes where the underlying classical pattern of the mesh pattern is \(21\).

\begin{proposition}[First Dominating Pattern Rule]
    \label{prop:dom1}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 1\), then \(m_1 \coinc{\pi} m_2 \)
    if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
    \end{enumerate}
\end{proposition}

This rule can be understood in graphical form. In the pattern in
\FigureRef{fig:rule1} we can gain shading in the boxes \(\boks{0,2}\), \(\boks{2,0}\)
since if there is a point in either of these boxes there would be an occurrence of
the dominating pattern \(\perm{3,2,1}\).

@@

\begin{proof}[Proof of \PropositionRef{prop:dom1}]
    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} implies that
    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Now we consider a permutation \(\omega \in \av{\pi}\),
    containing an occurrence of \(m_1\). If there is a point in the
    region corresponding to the box \(\boks{a,b}\), then that point, along with the
    points of the occurrence of \(m_1\), form an occurrence of \(\sigma^\boks{a,b}\).
    Then condition~\eqref{prop:dom1:cont} of the
    proposition implies an occurrence of \(\pi\). Therefore there can be no
    points in this region, which implies the occurrence of \(m_1\) is an occurrence of \(m_2\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

@@

After implementing the First Dominating Pattern Rule we find that there are
\(29\) \(\coinc[(1)]{321}\)-classes, of mesh patterns where the underlying
classical pattern is \(21\). By Note~\NoteRef{note:main} there are
therefore exactly \(29\) \(\coinc{321}\)-classes.

\subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
By application of First Dominating Pattern Rule we obtain \(43\) \(\coinc[(1)]{231}\)-coincidence
classes between mesh patterns with \(21\) as an underlying classical pattern.
Experimentation shows that there are \(39 \coinc[comp]{231}\)-coincidence
classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns are in the same
\(\coinc[comp]{231}\)-class, but this is not explained by the First Dominating Pattern Rule.
\begin{equation*}
    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
\end{equation*}
